---
title: "CaseStudy2"
author: "Allen Miller"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Libraries & Set Environment
```{r}
library(ggplot2)
library(GGally)
library(caret)
library(MASS)
library(tidyverse)
library(e1071)
library(mvtnorm)
library(class)
library(plyr)
library(dplyr)
options(scipen = 999)
```

## Import Data
```{r}
attrition <- read.csv(file.choose())
#attrition <- read.csv("/Users/allenmiller/OneDrive - Southern Methodist University/Doing Data Science/CaseStudy2/CaseStudy2-data.csv")


attrition$Attrition <- factor(attrition$Attrition)
attrition$BusinessTravel <- factor(attrition$BusinessTravel)
attrition$Department <- factor(attrition$Department)
attrition$EducationField <- factor(attrition$EducationField)
attrition$Gender <- factor(attrition$Gender)
attrition$JobRole <- factor(attrition$JobRole)
attrition$MaritalStatus <- factor(attrition$MaritalStatus)
attrition$Over18 <- factor(attrition$Over18)
attrition$OverTime <- factor(attrition$OverTime)
attrition$Education <- factor(attrition$Education)
attrition$EnvironmentSatisfaction <- factor(attrition$EnvironmentSatisfaction)
attrition$JobInvolvement <- factor(attrition$JobInvolvement)
attrition$JobLevel <- factor(attrition$JobLevel)
attrition$JobSatisfaction <- factor(attrition$JobSatisfaction)
attrition$PerformanceRating <- factor(attrition$PerformanceRating)
attrition$RelationshipSatisfaction <- factor(attrition$RelationshipSatisfaction)
attrition$StockOptionLevel <- factor(attrition$StockOptionLevel)
attrition$WorkLifeBalance <- factor(attrition$WorkLifeBalance)

str(attrition)
```

#Explore the data
##Selected Variables
###Age(2), DistanceFromHome(7), JobLevel(16), MonthlyIncome(20), NumCompaniesWorked(22), OverTime(24)
```{r}
#Keep = Age, BusinessTravel, Department, DistanceFromHome (2,4,6,7)
attrition %>% dplyr::select(Attrition, Age, BusinessTravel, DailyRate, Department, DistanceFromHome)%>%
  ggpairs()

#Keep = Education, EnvironmentSatisfaction (8,12)
attrition %>% dplyr::select(Attrition, Education, EducationField, EnvironmentSatisfaction, Gender, HourlyRate)%>%
  ggpairs()

#Keep = JobInvolvement, JobLevel, JobRole JobSatisfaction (15,16,17,18)
attrition %>% dplyr::select(Attrition, JobInvolvement, JobLevel, JobRole, JobSatisfaction, MaritalStatus)%>%
  ggpairs()

#Keep = MonthlyIncome, NumCompaniesWorked, OverTime (20,22,24)
attrition %>% dplyr::select(Attrition, MonthlyIncome, MonthlyRate, NumCompaniesWorked, Over18, OverTime)%>%
  ggpairs()

#Keep = PerformanceRating, RelationshipSatisfaction, StockOptionLevel, TotalWorkingYears (26,27,29,30)
attrition %>% dplyr::select(Attrition, PercentSalaryHike, PerformanceRating, RelationshipSatisfaction, StockOptionLevel, TotalWorkingYears)%>%
  ggpairs()

#Keep = WorkLifeBalance, YearsAtCompany, YearsInCurrentRole (32,33,34)
attrition %>% dplyr::select(Attrition, TrainingTimesLastYear, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole)%>%
  ggpairs()

#Keep = YearsSinceLastPromotion, YearsWithCurrManager (35,36)
attrition %>% dplyr::select(Attrition, YearsSinceLastPromotion, YearsWithCurrManager)%>%
  ggpairs()

```

#Run Neive Bayes with 22 selected variables
##Results
###Mean Accuracy = 83.25%
###Mean Sensitivity = 87.75%
###Mean Specificity = 61.03%
```{r}

set.seed(10)

iterations = 10

masterAcc = matrix(nrow = iterations)
masterSen = matrix(nrow = iterations)
masterSpec = matrix(nrow = iterations)

splitPerc = .7 #Training / Test split Percentage

for(j in 1:iterations)
{
  trainIndices = sample(1:dim(attrition)[1],round(splitPerc * dim(attrition)[1]))
  train = attrition[trainIndices,]
  test = attrition[-trainIndices,]
  
  model = naiveBayes(train,train$Attrition, laplace = 1)
  table(predict(model,test[,c(2,4,6,7,8,12,15,16,17,18,20,22,24,26,27,29,30,32,33,34,35,36)]),test$Attrition)
  CMB = confusionMatrix(table(predict(model,test[,c(2,4,6,7,8,12,15,16,17,18,20,22,24,26,27,29,30,32,33,34,35,36)]),test$Attrition))
  masterAcc[j] = CMB$overall[1]
  masterSen[j] = CMB$byClass[1]
  masterSpec[j] = CMB$byClass[2]
}
MeanAcc = colMeans(masterAcc)
MeanSen = colMeans(masterSen)
MeanSpec = colMeans(masterSpec)

#Results
MeanAcc
MeanSen
MeanSpec
```

#Use Test Set inside the Model
```{r}
#Blank DataSet
attritionTest <- read.csv(file.choose())
#attritionTest <- read.csv("/Users/allenmiller/OneDrive - Southern Methodist University/Doing Data Science/CaseStudy2/CaseStudy2CompSet No Attrition.csv")

summary(attritionTest)

#Change Test Data to Factors where applicable
attritionTest$BusinessTravel <- factor(attritionTest$BusinessTravel)
attritionTest$Department <- factor(attritionTest$Department)
attritionTest$EducationField <- factor(attritionTest$EducationField)
attritionTest$Gender <- factor(attritionTest$Gender)
attritionTest$JobRole <- factor(attritionTest$JobRole)
attritionTest$MaritalStatus <- factor(attritionTest$MaritalStatus)
attritionTest$Over18 <- factor(attritionTest$Over18)
attritionTest$OverTime <- factor(attritionTest$OverTime)
attritionTest$Education <- factor(attritionTest$Education)
attritionTest$EnvironmentSatisfaction <- factor(attritionTest$EnvironmentSatisfaction)
attritionTest$JobInvolvement <- factor(attritionTest$JobInvolvement)
attritionTest$JobLevel <- factor(attritionTest$JobLevel)
attritionTest$JobSatisfaction <- factor(attritionTest$JobSatisfaction)
attritionTest$PerformanceRating <- factor(attritionTest$PerformanceRating)
attritionTest$RelationshipSatisfaction <- factor(attritionTest$RelationshipSatisfaction)
attritionTest$StockOptionLevel <- factor(attritionTest$StockOptionLevel)
attritionTest$WorkLifeBalance <- factor(attritionTest$WorkLifeBalance)

str(attritionTest)

#Run Model and Predict On Test Data
model1 = naiveBayes(attrition[,c(2,4,6,7,8,12,15,16,17,18,20,22,24,26,27,29,30,32,33,34,35,36)],attrition$Attrition, laplace = 1)
attritionPredict <- unlist(predict(model1,attritionTest[,c(3,5,6,7,11,14,15,16,17,19,21,23,25,26,28,29,31,32,33,34,35)]))
attritionTest$Attrition <- attritionPredict

summary(attritionTest)

#Create a Data Frame with only ID and Attrition and Sort by ID
attritionFinal <- attritionTest %>% dplyr::select(ID, Attrition)
attritionFinal <- attritionFinal[order(attritionFinal$ID),]

#verify data and sorting
attritionFinal

#Save to CSV
#write.csv(attritionFinal,"Case2PredictionsClassifyMiller.csv",row.names = FALSE)

```

#Farther Data Analysis
```{r}
#Look at how long employees have been at their company by Job Role
##Managers have the highest median time at the company as well as the largest range of time
###Sales Reps typically spend the shortest amount of time at the company with the Median time well below 5 years and 75% of all employees spending less than 5 years at the company
attrition %>% ggplot(aes(x = JobRole, y = YearsAtCompany, color = JobRole)) + geom_boxplot() +
  labs(title = "Job Role vs Years at Company", x = "Job Role", y = "Years at Company" , color = "Job Role")

#Looking at how Years At Company and Marital Status influence Business Travel
##Divorced employees who travel have a higher median time at the company
###Single Employees all have a median time at company around 5 regardless of the amount of business travel
####All employees who travel frequently have a higher median time at the company regardless of Marital Status
attrition %>% ggplot(aes(x = BusinessTravel, y = YearsAtCompany, color = MaritalStatus)) + geom_boxplot() +
  labs(title = "Business Travel vs Years at Company colored by Marital Status", x = "Business Travel",
       y = "Years At Company", color = "Marital Status")

#How Age and Gender influence Monthly income
##Seems to be some positive correlation between Monthly Income and Age though it does not appear to be linear
###The distribution of Monthly Income is very similar between Male and Female
attrition %>% ggplot(aes(x = Age, y = MonthlyIncome, color = Gender)) + geom_point() + facet_wrap(~ Gender) +
  labs(title = "How Age and Gender influence Monthly Income", x = "Age", y = "Monthly Income", color = "Gender")

#Look at Mean Monthly Income by Job Role
##Managers have the highest Mean Income while Sales Reps have the lowest Mean Income
###33% of the Job Roles have a Mean Income over 7,500
####A little more than 50% of the Job Roles have a Mean Income over 5,000
#####There are 4 Job Roles with a Mean Income of around 2,500 or less
AvgRoleIncome <- ddply(attrition, "JobRole", summarise, MMI = mean(MonthlyIncome)) 
ggplot(AvgRoleIncome, aes(x = JobRole, y = MMI, fill = JobRole)) + geom_bar(stat = "identity") +
  labs(title = "Mean Monthly Income per Job Role", x = "Job Role", y = "Mean Income (Monthly)", color = "Job Role")

#Histogram of Job Satisfaction in the 3 departments
##Human Resources has the fewest responses, however there is little variation in total responses for rating 1,3, and 4
###Overall most of the company is satisfied (rating 3 or 4) with their managers
attrition %>% ggplot(aes(x = RelationshipSatisfaction, fill = Department)) + geom_bar() + facet_grid(~Department) +
  labs(title = "Distribution of Relationship Satisfaction Rating of Manager by Department", x = "Satisfaction Rating",
       y = "Total Responses")

```

#Monthly Income Analysis
##Selecting the variables
```{r}
#Keep = YearsAtCompany, JobLevel, Age, OverTime
attrition %>% dplyr::select(MonthlyIncome, YearsAtCompany,JobLevel,Age,OverTime)%>%
  ggpairs()

#Keep = Department, Education
attrition %>% dplyr::select(MonthlyIncome, DailyRate,Department,Education,EducationField)%>%
  ggpairs()

#Keep = JobRole
attrition %>% dplyr::select(MonthlyIncome, Gender,HourlyRate,JobInvolvement,JobRole)%>%
  ggpairs()

#Keep = None
attrition %>% dplyr::select(MonthlyIncome, MonthlyRate,NumCompaniesWorked,PercentSalaryHike,PerformanceRating)%>%
  ggpairs()

#Keep = TotalWorkingYears
attrition %>% dplyr::select(MonthlyIncome, StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance)%>%
  ggpairs()

#Keep = YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager
attrition %>% dplyr::select(MonthlyIncome, YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)%>%
  ggpairs()

#Create a list to hold 11 chosen variables
custom.var = attrition %>% 
  dplyr::select(MonthlyIncome, ID, YearsAtCompany, JobLevel, Age, Department, Education, JobRole,
                OverTime, TotalWorkingYears, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)
```

#Create the Model and Test
##RMSE = 1018.206
```{r}
#Model Training
set.seed(11)
train.control <- trainControl(method = "cv", number = 10)
custom.model = train(MonthlyIncome~.-ID, data=custom.var,
                   method="lmStepAIC",
                   trControl = train.control,
                   trace=FALSE)
custom.model$results
custom.model$finalModel
summary(custom.model$finalModel)
print(custom.model)


#Load in Test Data and Clean Data to match train data
MonthlyIncomeTest <- read.csv(file.choose())
#MonthlyIncomeTest <- read.csv("/Users/allenmiller/OneDrive - Southern Methodist University/Doing Data Science/CaseStudy2/CaseStudy2CompSet No Salary.csv")

str(MonthlyIncomeTest)

MonthlyIncomeTest$Attrition <- factor(MonthlyIncomeTest$Attrition)
MonthlyIncomeTest$BusinessTravel <- factor(MonthlyIncomeTest$BusinessTravel)
MonthlyIncomeTest$Department <- factor(MonthlyIncomeTest$Department)
MonthlyIncomeTest$EducationField <- factor(MonthlyIncomeTest$EducationField)
MonthlyIncomeTest$Gender <- factor(MonthlyIncomeTest$Gender)
MonthlyIncomeTest$JobRole <- factor(MonthlyIncomeTest$JobRole)
MonthlyIncomeTest$MaritalStatus <- factor(MonthlyIncomeTest$MaritalStatus)
MonthlyIncomeTest$Over18 <- factor(MonthlyIncomeTest$Over18)
MonthlyIncomeTest$OverTime <- factor(MonthlyIncomeTest$OverTime)

str(MonthlyIncomeTest)

#Create a list to hold 11 chosen variables
IncomeTestData = attrition %>% 
  dplyr::select(MonthlyIncome, ID, YearsAtCompany, JobLevel, Age, Department, Education, JobRole,
                OverTime, TotalWorkingYears, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)

#Test Data and Add it to a sorted data frame by ID
IncomeTestData$MonthlyIncome <- predict(custom.model, IncomeTestData)
custom.result <- IncomeTestData %>% dplyr::select(ID,MonthlyIncome)
custom.result <- custom.result[order(custom.result$ID),]

#Verify Order and View Predictions
custom.result

#Save CSV File
#write.csv(custom.result,"Case2PredictionsRegressMiller.csv",row.names = FALSE)
```





















